---
title: "Furniture"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Furniture}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(furniture)
```

## Using `furniture`

We will first make a ficticious data set:
```{r data}
df <- data.frame(a = rnorm(100, 1.5, 2), 
                 b = seq(1, 100, 1), 
                 c = c(rep("control", 40), rep("Other", 7), rep("treatment", 50), rep("None", 3)),
                 d = c(sample(1:1000, 90, replace=TRUE), rep(-99, 10)))
```

There are two functions that we'll demonstrate here:

1. `washer`
2. `table1`

## Washer

`washer` is a great function for quick data cleaning. In situations where there are placeholders, extra levels in a factor, or several values need to be changed to another.
```{r washer}
library(tidyverse)

df <- df %>%
  mutate(d = washer(d, -99),  ## changes the placeholder -99 to NA
         c = washer(c, "Other", "None", value = "control")) ## changes "Other" and "None" to "Control"
```


## Table1

Now that the data is "washed" we can start exploring and reporting.
```{r table1}
table1(df, a, b, factor(c), d)
```

The variables must be numeric or factor. Since we use a special type of evaluation (i.e. Non-Standard Evaluation) we can change the variables in the function (e.g., `factor(c)`). This can be extended to making a whole new variable in the function as well.

```{r table1.2}
table1(df, a, b, d, ifelse(a > 1, 1, 0))
```

This is just the beginning though. Two powerful things the function can do are shown below:

```{r table1.3}
table1(df, a, b, d, ifelse(a > 1, 1, 0),
       splitby=~factor(c), 
       test=TRUE)
```

The `splitby = ~factor(c)` stratifies the means and counts by a factor variable (in this case either control or treatment). When we use this we can also automatically compute tests of significance using `test=TRUE`. 

Finally, you can polish it quite a bit using a few other options. For example, you can do the following:
```{r table1.4}
table1(df, a, b, d, ifelse(a > 1, 1, 0),
       splitby=~factor(c), 
       test=TRUE,
       var_names = c("A", "B", "D", "New Var"),
       splitby_labels = c("Control", "Treatment"))
```

You can also format the numbers (adding a comma for big numbers such as in 20,000 instead of 20000):
```{r table1.5}
table1(df, a, b, d, ifelse(a > 1, 1, 0),
       splitby=~factor(c), 
       test=TRUE,
       var_names = c("A", "B", "D", "New Var"),
       splitby_labels = c("Control", "Treatment"),
       format_number = TRUE)
```

This can also be outputted as a latex, markdown, or pandoc table (matching all the output types of `knitr::kable`). Below shows how to do a latex table:
```{r table1.6}
table1(df, a, b, d, ifelse(a > 1, 1, 0),
       splitby=~factor(c), 
       test=TRUE,
       var_names = c("A", "B", "D", "New Var"),
       splitby_labels = c("Control", "Treatment"),
       output_type = "latex")
```

## Simple Table 1

Last item to show you: `simple_table1()`. This instead of reporting counts and percentages for categorical variables, it reports only percentages. I have more plans for this function but for now, that is the main benefit over `table1()`.

```{r simple_table1.1}
simple_table1(df, a, b, d, ifelse(a > 1, 1, 0),
              splitby=~factor(c), 
              test=TRUE,
              var_names = c("A", "B", "D", "New Var"),
              splitby_labels = c("Control", "Treatment"))
```

## Conclusion

The three functions: `table1`, `simple_table1` and `washer` add simplicity to cleaning up and understanding your data. Use these pieces of furniture to make your quantitative life a bit easier.

