---
title: "Table 1"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Furniture}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
date: "`r Sys.Date()`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(furniture)
```

# Making Table 1

This vignette demonstrates the main function of the `furniture` package--`table1`. The main parts of the package are below:

```{r structure, eval=FALSE}
table1(.data, ..., splitby, test, output_type, format_output, format_number, NAkeep, piping)
```

It contains several useful features for summarizing your data:

1. It simply summarizes many variables succinctly providing means/counts and SD's/percentages.
2. The summary can be by a grouping factor (i.e., splitby).
3. It uses a similar API to the popular tidyverse groups of packages.
4. It can be used in piping.
5. It can give bivariate test results for the variable with the grouping variable, which provides the correct test type depending on the variable types.
6. It is flexible as to its output: can be printed in regular console output or it can be printed in latex, markdown, and pandoc (see `knitr::kable`).
7. Numbers can be formatted nicely.

## 

```{r data}
## Create Ficticious Data containing several types of variables
df <- data.frame(a = rnorm(1000),
                 b = runif(1000),
                 c = factor(sample(c(1,2,3,4), 1000, replace=TRUE)),
                 d = factor(sample(c(0,1), 1000, replace=TRUE)),
                 e = trunc(rnorm(1000, 20, 5)))
```